Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-03-03 article-models-tt.tex
motivating property made into lemma
compatibility of comp with ctxext
a start for issue #27

Using the rules of the compatibility of substitution with weakening and of the
compatibility of weakening with itself, we see that we can show
\begin{equation*}
\jfameq{{\Gamma}{A}}{\subst{f}{\ctxwk{A}{{B}{C}}}}{\ctxwk{A}{C}}
\end{equation*}
for any three families $A$, $B$ and $C$ in context $\Gamma$ and any $\jterm{{\Gamma}{A}}{\ctxwk{A}{B}}{f}$.
It follows that for $\jterm{{\Gamma}{B}}{\ctxwk{B}{C}}{g}$ we can compose $f$
with $g$ to obtain a term of $\ctxwk{A}{C}$ in context $\ctxext{\Gamma}{A}$.
In the following definition, we work with in a slightly greater generality.

@@

>>>>

Using the rules of the compatibility of substitution with weakening and of the
compatibility of weakening with itself, we see that we can show

\begin{lem}
The inference rule
\begin{equation*}
\inference
  {\jfam{\Gamma}{A}
   \jfam{\Gamma}{B}
   \jfam{\Gamma}{C}
   \jhom{\Gamma}{A}{B}{f}}
  {\jfameq{{\Gamma}{A}}
    {\subst{f}{\ctxwk{A}{{B}{C}}}}
    {\ctxwk{A}{C}}}
\end{equation*}
is valid.
\end{lem}

\begin{proof}
Let $\jfam{\Gamma}{A}$, $\jfam{\Gamma}{B}$, $\jfam{\Gamma}{C}$ and $\jhom{\Gamma}{A}{B}{f}$.
Then we have the judgmental equalities
\begin{align*}
\subst{f}{\ctxwk{A}{{B}{C}}}
& \jdeq \subst{f}{\ctxwk{{A}{B}}{{A}{C}}}\\
& \jdeq \ctxwk{A}{C}.
\end{align*}
\end{proof}

It follows that for $\jterm{{\Gamma}{B}}{\ctxwk{B}{C}}{g}$ we can compose $f$
with $g$ to obtain a term of $\ctxwk{A}{C}$ in context $\ctxext{\Gamma}{A}$.
In the following definition, we work with in a slightly greater generality.

@@

We also have the following lemma about the compatibility of pairing and composition:

\begin{lem}
The following inference rule is valid
\begin{align*}
& \inference
  {\jhom{\Gamma}{A}{B}{f}
   \jhom{\Gamma}{B}{C}{g}
   \jfam{{\Gamma}{C}}{R}
   \jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}}
  {\jhomeq{\Gamma}{A}{{C}{R}}
    {\jcomp{f}{\ctxext{g}{w}}}
    {\ctxext{\jcomp{f}{g}}{\jcomp{f}{w}}}}
\end{align*}
\end{lem}

\begin{proof}
Let $\jhom{\Gamma}{A}{B}{f}$, $\jhom{\Gamma}{B}{C}{g}$, $\jfam{{\Gamma}{C}}{R}$
and $\jterm{{\Gamma}{B}}{\subst{g}{\ctxwk{B}{R}}}{w}$. Then we have the
judgmental equalities
\begin{align*}
\jcomp{f}{\ctxext{g}{w}}
& \jdeq \subst{f}{\ctxwk{A}{\ctxext{g}{w}}}\\
& \jdeq \subst{f}{\ctxext{\ctxwk{A}{g}}{\ctxwk{A}{w}}}\\
& \jdeq \ctxext{\subst{f}{\ctxwk{A}{g}}}{\subst{f}{\ctxwk{A}{w}}}\\
& \jdeq \ctxext{\jcomp{f}{g}}{\jcomp{f}{w}}.
\end{align*}
\end{proof}

\subsection{The unit type in structural type theory}
In this section we explore what we get if we pose compatibility conditions on
an inductively defined unit type. We will assume that not only the unit type
is compatible with extension, weakening, substitution and identity functions,
but also its induction principle should be compatible with those.
